---
title: "DDM"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
urlcolor: blue
date: "2023-07-21"
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>


```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,RWiener, tidybayes, posterior, furrr,gganimate, cmdstanr)

```

Lets start off making a function that simulates the weiner process:

```{r}
set.seed(1111)

make_weiner = function(data){
  
timesteps = data$max_time
w0 = data$bias
sd = data$sd

bound = data$bound

drift = data$drift
#keeping track of position
w = array(NA,timesteps+1)
w[1] = 0

for(i in 1:timesteps){
    if(w[i] > bound){
      w[i:(i+5)] = bound
      
      break
    }else if(w[i] < -bound){
      w[i:(i+5)] = -bound
      break
    }
    w[i+1] = w[i]+rnorm(1,0,sd)+drift
}

q = data.frame(x = 1:timesteps,
           y = w[1:timesteps],
           bound = bound,
           bias = bias,
           sd = sd,
           max_time = max_time,
           drift = drift,
           col = round(rnorm(1,100,1000),0)) #this is id
q = q %>% mutate(resp = ifelse(y == bound, "upper","lower"))

return(list(q))
}
```

Next we then simulate 5 such process!

```{r}
max_time = seq(500,length.out = 1)

bias = seq(0.5,length.out = 1)

sd = seq(0.1,length.out = 1)

bound = seq(1,2,length.out = 1)

drift = seq(0,length.out = 1)

replicate = 1:5


params = expand.grid(max_time = max_time,
                     bias = bias,
                     sd = sd,
                     bound = bound,
                     drift = drift,
                     replicate = replicate) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(params, params$id)

plan(multisession, workers = 3)
weiners = future_map(data_list, ~make_weiner(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

data = map_dfr(weiners,1)

q = na.omit(data) %>% 
  ggplot(aes(x = x, y = y, color = as.factor(col)))+
  xlab("Time")+
  ylab("x")+
  coord_cartesian(ylim = c(min(na.omit(data)$y),max(na.omit(data)$y)))+
  geom_line(linewidth=1)+
  theme_classic()+
  geom_hline(yintercept = bound, linetype = 2)+
  geom_hline(yintercept = -bound, linetype = 2)+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  #transition_time(x)+
  #shadow_wake(wake_length = 0.1, alpha = FALSE)+
  #shadow_mark(alpha = 0.3, size = 0.5)+
  transition_reveal(x)

q
```

We can now play around with different drift rates and animate those!

```{r differing driftrates}
max_time = seq(500,length.out = 1)

bias = seq(0.5,length.out = 1)

sd = seq(1,length.out = 1)

bound = seq(1,2,length.out = 1)

drift = seq(-1,1,length.out = 5)

replicate = 1:5


params = expand.grid(max_time = max_time,
                     bias = bias,
                     sd = sd,
                     bound = bound,
                     drift = drift,
                     replicate = replicate) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(params, params$id)

plan(multisession, workers = 3)
weiners = future_map(data_list, ~make_weiner(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

data = map_dfr(weiners,1)

q = na.omit(data) %>% 
  ggplot(aes(x = x, y = y, color = as.factor(col)))+
  xlab("Time")+
  ylab("x")+
  facet_wrap(~drift, nrow = 3, ncol = 2, labeller = label_both)+
  coord_cartesian(ylim = c(min(na.omit(data)$y),max(na.omit(data)$y)))+
  geom_line(linewidth=1)+
  theme_classic()+
  geom_hline(yintercept = bound, linetype = 2)+
  geom_hline(yintercept = -bound, linetype = 2)+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  #transition_time(x)+
  #shadow_wake(wake_length = 0.1, alpha = FALSE)+
  #shadow_mark(alpha = 0.3, size = 0.5)+
  transition_reveal(x)

q
```

Lastly we can look at the RT distribution for both lower and upper bound hits for different drift rates:

```{r}
max_time = seq(500,length.out = 1)

bias = seq(0.5,length.out = 1)

sd = seq(1,length.out = 1)

bound = seq(1,2,length.out = 1)

drift = seq(-1,1,length.out = 5)

replicate = 1:100


params = expand.grid(max_time = max_time,
                     bias = bias,
                     sd = sd,
                     bound = bound,
                     drift = drift,
                     replicate = replicate) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(params, params$id)


plan(multisession, workers = 3)
weiners = future_map(data_list, ~make_weiner(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

data = map_dfr(weiners,1)

data %>% na.omit(data) %>% group_by(col,drift) %>% summarize(last(resp))

qq = na.omit(data) %>% group_by(col,drift) %>% summarize(x = max(x), resp = last(resp))  %>% ungroup() %>% 
  ggplot(aes(x = x)) + 
  geom_histogram(aes(fill = resp,bins = 30), position = position_dodge())+
  theme_classic()+
  facet_wrap(~drift, labeller = label_both)
qq

```

```{r}

df = data.frame()
for(d in drift){
  df = rbind(df,RWiener::rwiener(1000,alpha = bound,tau = 0.01, beta = bias, delta = d)%>% mutate(drift = d))
}


qq = na.omit(df) %>% 
  mutate(upper = ifelse(resp == "upper", T,F)) %>% 
  ggplot(aes(x = q)) + 
  geom_histogram(aes(fill = upper,bins = 30), position = position_dodge())+
  theme_classic()+
  facet_wrap(~drift, labeller = label_both)


qq
```

There does seem to be a difference in how the RWeiner function uses the driftrate and how i use it? Why??
