---
title: "ERLDMM"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
urlcolor: blue
date: "2023-07-21"
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>


```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,RWiener, tidybayes, posterior, furrr,gganimate)

```

Lets start off making a function that simulates the weiner process:

```{r}
make_weiner = function(data){
  
timesteps = data$max_time
w0 = data$bias
sd = data$sd

bound = data$bound

drift = data$drift

w = array(NA,timesteps+1)
w[1] = 0
for(i in 1:timesteps){
    if(w[i] > bound){
      w[i:(i+5)] = bound
      
      break
    }else if(w[i] < -bound){
      w[i:(i+5)] = -bound
      break
    }
    w[i+1] = w[i]+rnorm(1,0,sd)+drift
  }
q = data.frame(x = 1:timesteps,
           y = w[1:timesteps],
           bound = bound,
           bias = bias,
           sd = sd,
           max_time = max_time,
           drift = drift,
           col = round(rnorm(1,100,1000),0))

return(list(q))
}
```

Next we then simulate 5 such process!

```{r}
max_time = seq(500,length.out = 1)

bias = seq(0.5,length.out = 1)

sd = seq(0.1,length.out = 1)

bound = seq(1,2,length.out = 1)

drift = seq(0,length.out = 1)

replicate = 1:5


params = expand.grid(max_time = max_time,
                     bias = bias,
                     sd = sd,
                     bound = bound,
                     drift = drift,
                     replicate = replicate) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(params, params$id)

plan(multisession, workers = availableCores()-3)
weiners = future_map(data_list, ~make_weiner(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

data = map_dfr(weiners,1)

q = na.omit(data) %>% 
  ggplot(aes(x = x, y = y, color = as.factor(col)))+
  xlab("Time")+
  ylab("x")+
  coord_cartesian(ylim = c(min(na.omit(data)$y),max(na.omit(data)$y)))+
  geom_line(linewidth=1)+
  theme_classic()+
  geom_hline(yintercept = bound, linetype = 2)+
  geom_hline(yintercept = -bound, linetype = 2)+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  #transition_time(x)+
  #shadow_wake(wake_length = 0.1, alpha = FALSE)+
  #shadow_mark(alpha = 0.3, size = 0.5)+
  transition_reveal(x)

q
```

We can now play around with different drift rates and animate those!

```{r differing driftrates}
max_time = seq(500,length.out = 1)

bias = seq(0.5,length.out = 1)

sd = seq(1,length.out = 1)

bound = seq(1,2,length.out = 1)

drift = seq(-1,1,length.out = 5)

replicate = 1:5


params = expand.grid(max_time = max_time,
                     bias = bias,
                     sd = sd,
                     bound = bound,
                     drift = drift,
                     replicate = replicate) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(params, params$id)

plan(multisession, workers = availableCores()-3)
weiners = future_map(data_list, ~make_weiner(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

data = map_dfr(weiners,1)

q = na.omit(data) %>% 
  ggplot(aes(x = x, y = y, color = as.factor(col)))+
  xlab("Time")+
  ylab("x")+
  facet_wrap(~drift, nrow = 3, ncol = 2, labeller = label_both)+
  coord_cartesian(ylim = c(min(na.omit(data)$y),max(na.omit(data)$y)))+
  geom_line(linewidth=1)+
  theme_classic()+
  geom_hline(yintercept = bound, linetype = 2)+
  geom_hline(yintercept = -bound, linetype = 2)+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  #transition_time(x)+
  #shadow_wake(wake_length = 0.1, alpha = FALSE)+
  #shadow_mark(alpha = 0.3, size = 0.5)+
  transition_reveal(x)

q
```

Lastly we can look at the RT distribution for both lower and upper bound hits for different drift rates:

```{r}
max_time = seq(500,length.out = 1)

bias = seq(0.5,length.out = 1)

sd = seq(1,length.out = 1)

bound = seq(1,2,length.out = 1)

drift = seq(-1,1,length.out = 5)

replicate = 1:1000


params = expand.grid(max_time = max_time,
                     bias = bias,
                     sd = sd,
                     bound = bound,
                     drift = drift,
                     replicate = replicate) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(params, params$id)

plan(multisession, workers = availableCores()-3)
weiners = future_map(data_list, ~make_weiner(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

data = map_dfr(weiners,1)


qq = na.omit(data) %>% group_by(col,drift) %>% summarize(x = max(x)) %>% 
  mutate(upper = ifelse(col < 0, T,F)) %>% ungroup() %>% mutate(frame = 1:nrow(.)) %>% 
  ggplot(aes(x = x)) + 
  geom_histogram(aes(fill = upper,bins = 30), position = position_dodge())+
  theme_classic()+
  facet_wrap(~drift, labeller = label_both)
qq

```

```{r}

df = data.frame()
for(d in drift){
  df = rbind(df,RWiener::rwiener(1000,alpha = bound,tau = 0.01, beta = bias, delta = d)%>% mutate(drift = d))
}


qq = na.omit(df) %>% 
  mutate(upper = ifelse(resp == "upper", T,F)) %>% 
  ggplot(aes(x = q)) + 
  geom_histogram(aes(fill = upper,bins = 30), position = position_dodge())+
  theme_classic()+
  facet_wrap(~drift, labeller = label_both)


qq
```

Lets try and fit this in Stan!

```{r}
trials = 500
alpha = 2
delta = 0
beta = 0.5
tau = 0.1


parameters = data.frame(alpha,delta,beta,tau, trials)

data = rwiener(n = trials,
        alpha = alpha,
        delta = delta,
        beta = beta,
        tau = tau)

data_stan = list(Nu = nrow(data %>% filter(resp == "upper")),
                 Nl = nrow(data %>% filter(resp == "lower")),
                 RTu = data %>% filter(resp == "upper") %>% .$q,
                 RTl = data %>% filter(resp == "lower") %>% .$q,
                 minRT = min(data$q),
                 run_estimation = 1)


mod = cmdstanr::cmdstan_model(here::here("HDDM.stan"))


fit <- mod$sample(
    data = data_stan,
    chains = 4,
    parallel_chains = 4,
    adapt_delta = 0.9,
    max_treedepth = 12)



rt_pattern <- "out\\[\\d+,1\\]"
choice_pattern <- "out\\[\\d+,2\\]"


Rts = fit$summary() %>%
  filter(grepl(rt_pattern, variable))


Rts = as_draws_df(fit$draws()) %>%
  select(matches(rt_pattern))

Rts %>% slice(sample(1:4000,10)) %>% pivot_longer(everything()) %>% ggplot(aes(x = value))+geom_histogram()

choice = fit$summary() %>%
  filter(grepl(choice_pattern, variable))

choice = as_draws_df(fit$draws()) %>%
  select(matches(choice_pattern))


posteriors = as_draws_df(fit) %>% dplyr::select(any_of(names(parameters))) %>% mutate(prior = F)
priors = as_draws_df(fit) %>% dplyr::select(starts_with("prior_")) %>% rename_with(~gsub("^prior_", "", .), everything()) %>% mutate(prior = T)

rbind(posteriors,priors) %>% 
  pivot_longer(cols = -prior) %>% 
  ggplot(aes(x = value, fill = prior))+
  geom_density(alpha = 0.5)+
  theme_classic()+
  facet_wrap(~name, scales = "free")+
  geom_vline(data = parameters %>% select(-trials) %>% pivot_longer(everything()), aes(xintercept = value))


draws = 10
#pp_check
as_draws_df(fit$draws()) %>%
  select(matches(rt_pattern)) %>% slice(sample(1:4000,draws)) %>% 
  pivot_longer(everything()) %>% mutate(estimated = TRUE, slice = rep(1:trials,draws)) %>% 
  ggplot()+
  geom_density(aes(x = value, group = slice), color = "lightblue")+
  geom_density(data = data %>% mutate(estimated = FALSE), aes(x = q), color = "red")+
  theme_classic()
```


# Lets do some parameter recovery on this model!

## fit model
```{r}
fit_model = function(parameters){
  id = parameters$id
  
  data = rwiener(n = parameters$trials,
          alpha = parameters$alpha,
          delta = parameters$delta,
          beta = parameters$beta,
          tau = parameters$tau)
  
  data_stan = list(Nu = nrow(data %>% filter(resp == "upper")),
                   Nl = nrow(data %>% filter(resp == "lower")),
                   RTu = data %>% filter(resp == "upper") %>% .$q,
                   RTl = data %>% filter(resp == "lower") %>% .$q,
                   minRT = min(data$q),
                   RTbound = 0)
  
  
  mod = cmdstanr::cmdstan_model(here::here("HDDM.stan"))
  
  
  fit <- mod$sample(
      data = data_stan,
      seed = seed,
      chains = 4,
      parallel_chains = 4,
      adapt_delta = 0.9,
      max_treedepth = 12)
  
  
  
  posteriors = as_draws_df(fit$summary()) %>% dplyr::filter(variable %in% names(parameters))
  diag = data.frame(fit$diagnostic_summary(), id)
  
  data = posteriors %>% mutate(num_div = diag$num_divergent,
                               tree_depth = diag$num_max_treedepth,
                               real_alpha = parameters$alpha,
                               real_delta = parameters$delta,
                               real_beta = parameters$beta,
                               real_tau = parameters$tau,
                               trials = parameters$trials,
                               id = id) %>% select(-contains("."))
  return(list(data, diag))
}
```

```{r}
trials = seq(10,100,by = 10)
alpha = seq(0,5,by = 1)
delta = seq(-4,4,by = 1)
beta = seq(0.1,0.9,by = 0.1)
tau = seq(0.1,0.4,by = 0.1)

parameters = expand.grid(alpha = alpha,delta = delta,beta = beta,tau = tau, trials = trials) %>% 
  mutate(id = 1:nrow(.))

data_list <- split(parameters, parameters$id)
```


```{r}
#cores = availableCores()-1

plan(multisession, workers = cores)

possfit_model = possibly(.f = fit_model, otherwise = "Error")

results <- future_map(data_list, ~possfit_model(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))

error_indices <- which(results == "Error")

unique(error_indices)

results2 = results[results != "Error"]
```


lets look at the divergences
```{r}
results2[[2]]

divergence = map_dfr(results2, 2)

divergence %>% median_qi(num_divergent)
```
there are none which is good now at the parameter values

```{r, fig.height=10, fig.width=10}
params = map_dfr(results2, 1)

params %>% filter(variable == "alpha") %>% 
  ggplot(aes(x = mean, fill = as.factor(real_alpha)))+
  geom_density(alpha = 0.5)+
  theme_classic()+
  geom_vline(aes(xintercept = real_alpha))+
  facet_wrap(~trials)

```


